---
title: "ws"
output: html_document
date: "2025-02-06"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# TODO:
https://docs.google.com/document/d/1IYLTEWNzi34iagvPAwtEt4nIZdpzWqoQ5zt4LkySbhg/edit?tab=t.0 
1. merge 2021-2023 data with original exam data 
2. extract my variables 2013-2023 and demographics variables with SEQN
3. recode variables based on excel logic, document everything
4. save RData and uplaod to github

```{r}
# load packages here
library(haven)
library(dplyr)
source('/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/R/funcs.R')
```

# Load 1999-2020 exam data (nhanesA format)
```{r}
rm(list = ls())
load('/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/Examinations.RData')
```

# Save as dataframe with 1999-2020 data
```{r}
exnames <- c("AUX", "BPX", "BMX", 'DXX', 'BPXO', 'OHXDENT')

ex_dat <- ex_dat[names(ex_dat) %in% exnames]

ex_df <- combine_surveys(ex_dat)

save(ex_df, file = "/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/Examinations_1999_to_2020.RData")
```

# Load 2021-2023 data for exam and demographics
```{r}
rm(list = ls())

load('/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/Examinations_1999_to_2020.RData')
load('/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/Demographics.RData')
demo_df <- demo_dat$DEMO

BAX_L <- read_xpt("/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/2021_to_2023_data/exam/BAX_L.xpt")
BMX_L <- read_xpt("/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/2021_to_2023_data/exam/BMX_L.xpt")
BPXO_L <- read_xpt("/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/2021_to_2023_data/exam/BPXO_L.xpt")
LUX_L <- read_xpt("/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/2021_to_2023_data/exam/LUX_L.xpt")
DEMO_L <- read_xpt("/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/2021_to_2023_data/demographics/DEMO_L.xpt")
```

# Merge to create 1999-2023 data
```{r}
# Add Year = 2021 to all smaller datasets
datasets <- list(BAX_L, BMX_L, BPXO_L, LUX_L)

datasets <- lapply(datasets, function(df) {
  df <- as.data.frame(df)  # Ensure it's a data frame
  df <- df %>% mutate(Year = '2021')  # Add Year
  return(df)
})

# Perform full joins across all datasets
full_joined_data <- Reduce(function(x, y) full_join(x, y, by = c("SEQN", "Year")), datasets)

# Bind ex_df with the fully joined dataset
ex_df_joined <- bind_rows(ex_df, full_joined_data)


# Merge Demo data
datasets <- list(DEMO_L)
datasets <- lapply(datasets, function(df) {
  df <- as.data.frame(df)  # Ensure it's a data frame
  df <- df %>% mutate(Year = '2021')  # Add Year
  return(df)
})
demo_df_joined <- bind_rows(demo_df, datasets)
duplicate_seqn <- demo_df_joined %>%
  group_by(SEQN, Year) %>%
  filter(n() > 1)  # Keep only duplicates

# Display duplicate SEQN count
cat("Number of duplicated SEQN values:", nrow(duplicate_seqn), "\n")


# Saving 
demo_1999_to_2023 <- demo_df_joined
exam_1999_to_2023 <- ex_df_joined
save(demo_1999_to_2023, file = "/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/demo_1999_to_2023.RData")
save(exam_1999_to_2023, file = "/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/exam_1999_to_2023.RData")
```

# Extract necessary variables from the 1999-2023 data
```{r}
rm(list = ls())
load('/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/exam_1999_to_2023.RData')

exvars <- c(
  # ID, YEAR
  "SEQN", "Year",
  
  # BPX (Blood Pressure Exam)
  "BPQ150A", "BPQ150B", "BPQ150C", "BPQ150D", 
  "BPXPULS", "BPXPTY", "BPXML1", 
  "BPXSY1", "BPXDI1", "BPAEN1", "BPXSY2", "BPXDI2", "BPAEN2", 
  "BPXSY3", "BPXDI3", "BPAEN3", "BPXSY4", "BPXDI4", "BPAEN4",
  "BPXOSY1", "BPXODI1", 
  "BPXOSY2", "BPXODI2", 
  "BPXOSY3", "BPXODI3", 
  "BPXOPLS1", "BPXOPLS2", "BPXOPLS3",
  
  # BMX (Body Measurements Exam)
  "BMXBMI", "BMXWAIST", "BMXHIP",
  
  # DXX (Dual-energy X-ray Absorptiometry - DXA Scan)
  "DXXLSBMD", "DXDTRBMD", "DXDTRPF", "DXXLABMD", 
  "DXDLAPF", "DXXRABMD", "DXDRAPF",
  
  # AUX (Audiometry Exam)
  "AUAEXSTS", "AUQ020", "AUQ030", 
  "AUXU1K1R", "AUXU500R", "AUXU1K2R", "AUXU2KR", "AUXU3KR", 
  "AUXU4KR", "AUXU6KR", "AUXU8KR", "AUXU1K1L", "AUXU500L", 
  "AUXU1K2L", "AUXU2KL", "AUXU3KL", "AUXU4KL", "AUXU6KL", "AUXU8KL",
  "AUXR1K1R", "AUXR5CR", "AUXR1K2R", "AUXR2KR", "AUXR3KR", "AUXR4KR", 
  "AUXR6KR", "AUXR8KR", "AUXR1K1L", "AUXR5CL", "AUXR1K2L", "AUXR2KL", 
  "AUXR3KL", "AUXR4KL", "AUXR6KL", "AUXR8KL",
  
  # OHX (Oral Health Exam - Dental)
  "OHDEXSTS", "OHAREC", "OHAROCDT", "OHAROCGP", "OHAROCOH", 
  "OHAROCCI", "OHAROCDE",
  
  # BMDAVSAD (Bone Mineral Density - DXA Scan)
  "BMDAVSAD"
)

exam_subset <- exam_1999_to_2023 %>%
  filter(Year %in% c("2013", "2015", "2017", "P", "2021")) %>%
  select(any_of(exvars))  

missingness <- exam_subset %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(cols = everything(), names_to = "Column", values_to = "Missing_Percentage")
print(missingness)

```
# Data re-coding
```{r}
# Define BP variables for general years and 2021
bp_vars_general <- c("BPXSY1", "BPXDI1", "BPXSY2", "BPXDI2", "BPXSY3", "BPXDI3", "BPXSY4", "BPXDI4")
bp_vars_2021 <- c("BPXOSY1", "BPXODI1", "BPXOSY2", "BPXODI2", "BPXOSY3", "BPXODI3")
# Compute mean and standard deviation separately for 2021 and other years
bp_stats_general <- exam_subset %>%
  filter(Year != 2021) %>%
  summarise(across(all_of(bp_vars_general), list(mean = ~ mean(. , na.rm = TRUE),
                                                 sd = ~ sd(. , na.rm = TRUE)))) %>%
  pivot_longer(cols = everything(), names_to = c("Variable", ".value"), names_sep = "_") %>%
  mutate(Year = "Other Years")

bp_stats_2021 <- exam_subset %>%
  filter(Year == 2021) %>%
  summarise(across(all_of(bp_vars_2021), list(mean = ~ mean(. , na.rm = TRUE),
                                              sd = ~ sd(. , na.rm = TRUE)))) %>%
  pivot_longer(cols = everything(), names_to = c("Variable", ".value"), names_sep = "_") %>%
  mutate(Year = "2021")
# Combine both statistics
bp_stats_long <- bind_rows(bp_stats_general, bp_stats_2021)
# Display the results
print(bp_stats_long)


# Since not much fluctuation between BP1-BP3, we will take average of 1,2,3 to create new numerical variables called BPXSY and BPXDI
# Dual-Energy X-ray 
# select DXDTRBMD, DXXLABMD based on low correlation with other bmi variables and DXD 
exam_bp <- exam_subset %>%
  mutate(
    BPXSY = ifelse(Year == '2021', 
                   rowMeans(select(., BPXOSY1, BPXOSY2, BPXOSY3), na.rm = TRUE), 
                   rowMeans(select(., BPXSY1, BPXSY2, BPXSY3), na.rm = TRUE)),
    BPXDI = ifelse(Year == '2021', 
                   rowMeans(select(., BPXODI1, BPXODI2, BPXODI3), na.rm = TRUE), 
                   rowMeans(select(., BPXDI1, BPXDI2, BPXDI3), na.rm = TRUE))
  ) %>%
  select(SEQN, Year, BPXSY, BPXDI, BPXPULS, BMXBMI, DXDTRBMD, DXXLABMD)  # Keep only relevant columns

# removing BMXWAIST since high correlation with BMXBMI
print(cor.test(exam_subset$BMXBMI, exam_subset$BMXWAIST))

missingness <- exam_bp %>%
  summarise(across(everything(), ~ mean(is.na(.)) * 100)) %>%
  pivot_longer(cols = everything(), names_to = "Column", values_to = "Missing_Percentage")
print(missingness)

# Encode into hearing loss vs noraml based on literature's threshold for normal hearing 
exam_audio <- exam_subset %>%
  rowwise() %>%  
  mutate(hearing_loss = ifelse(
    any(c(AUXU1K1R, AUXU500R, AUXU1K2R, AUXU2KR, AUXU3KR, 
          AUXU4KR, AUXU6KR, AUXU8KR, AUXU1K1L, AUXU500L, 
          AUXU1K2L, AUXU2KL, AUXU3KL, AUXU4KL, AUXU6KL, AUXU8KL) > 21, 
        na.rm = TRUE), 
    "No",   # Normal
    "Yes"   # Hearing Loss
  )) %>%
  select(SEQN, Year, hearing_loss)
table(exam_audio$hearing_loss)


exam_2013_to_2023 <- exam_bp %>%
  left_join(exam_audio, by = c("SEQN", "Year"))

head(exam_2013_to_2023)
save(exam_2013_to_2023, file = '/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/2013_to_2023_cleaned/exam_2013_to_2023.RData')
```

# Sanity check
```{r}
rm(list = ls())
load('/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/2013_to_2023_cleaned/exam_2013_to_2023.RData')
load("/Users/taehyo/Library/CloudStorage/Dropbox/NYU/Research/Research/Code/ENAR-Risk-Factors-Analyses/data/cleaned/1999_to_2023_original/demo_1999_to_2023.RData")
```



